---
title: "Untitled"
author: "Dilpreet Singh"
date: "6/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 3- Leaflet

```{r include=FALSE}
# load packages
library(sf)
library(tidyverse)
library(ggplot2)
library(tmap)
library(tidygeocoder)
library(spdep)
library(sp)
library(leaflet)
library(leaflet.extras)
library(KernSmooth)
library(raster)
library(rgeos)
```

```{r}
# load final cleaned data for 2022
final_data_2022 <- read.csv("C:/Users/Dell/OneDrive/Desktop/capstone/final_data_cleaned_updated.csv")

# read ward boundaries shp file as vector
ward_boundaries <- st_read("Municipal Ward Boundaries_UTM83.shp")

#fetch crs of ward boundaries
the_crs <- st_crs(ward_boundaries) 

#convert the final_data_cleaned data frame to an sf object
final_data_sf <- st_as_sf(final_data_2022, coords = c("longitude","latitude"),crs=4326)

#transform crs of final_data_Sf to match crs of ward boundaries
 st_transform(final_data_sf, crs=the_crs) 

# transform crs to plot leaflet map (match projection of ward boundaries to leaflet)
ward_boundaries_1 <-
  st_transform(ward_boundaries, crs = "+proj=longlat +init=epsg:4326 +ellps=WGS84 +datum=WGS84 +no_defs")
```

# 2022 overall analysis

```{r}
# create basemap
basemap <- leaflet() %>% 
  # add different provider tiles
  addProviderTiles(
    "OpenStreetMap",
    # give the layer a name
    group = "OpenStreetMap"
  ) %>%
  addProviderTiles(
    "Stamen.Toner",
    group = "Stamen.Toner"
  ) %>%
  addProviderTiles(
    "Stamen.Terrain",
    group = "Stamen.Terrain"
  ) %>%
  addProviderTiles(
    "Esri.WorldStreetMap",
    group = "Esri.WorldStreetMap"
  ) %>%
  addProviderTiles(
    "Wikimedia",
    group = "Wikimedia"
  ) %>%
  addProviderTiles(
    "CartoDB.Positron",
    group = "CartoDB.Positron"
  ) %>%
  addProviderTiles(
    "Esri.WorldImagery",
    group = "Esri.WorldImagery"
  ) %>%
# add a layers control
  addLayersControl(
    baseGroups = c(
      "OpenStreetMap", "Stamen.Toner",
      "Stamen.Terrain", "Esri.WorldStreetMap",
      "Wikimedia", "CartoDB.Positron", "Esri.WorldImagery"
    ),
    # position it on the topleft
    position = "topleft"
  )
```

# map ward boundaries

```{r}
# Find a center point for each region (for labeling)
centers <-
  data.frame(gCentroid(as_Spatial(ward_boundaries_1), byid = TRUE))
centers$region <- c("9", "6", "7", "1", "2", "10", "3", "5", "4", "8")

# creating a custom color palette for coloring different wards
pal <-
  colorNumeric(palette = "Dark2",
               domain = as.numeric(ward_boundaries_1$NUMBER))
colorData <- as.numeric(ward_boundaries_1$NUMBER)

# ward boundaries map used for plotting contour plots
map_ward_boundaries <-
  leaflet() %>% addTiles() %>% 
  addPolygons(
    data = ward_boundaries_1,
    weight = 1,
    opacity = 1,
    fillOpacity = 0,
    smoothFactor = 0.5,
    color = 'blue'
  ) %>% # add labels for ward name
  addLabelOnlyMarkers(
    data = centers,
    lng = ~ x,
    lat = ~ y,
    label = ~ region,
    labelOptions = labelOptions(
      noHide = TRUE,
      direction = 'top',
      textOnly = TRUE,
      textsize = "17px"
    )
  )

# ward boundaries map with dark background
map_ward_boundaries_dark <-
  leaflet() %>% addProviderTiles("CartoDB.DarkMatter") %>%
  addPolygons(
    data = ward_boundaries_1,
    weight = 1,
    opacity = 0.5,
    fillOpacity = 0,
    smoothFactor = 0.5,
    color = 'white'
  ) %>% # add labels for ward name
  addLabelOnlyMarkers(
    data = centers,
    lng = ~ x,
    lat = ~ y,
    label = ~ region,
    labelOptions = labelOptions(
      noHide = TRUE,
      direction = 'top',
      textOnly = TRUE,
      textsize = "17px"
    )
  )

# add ward boundaries to base map along with ward labels and custom color pallet
map_1 <- basemap %>%
  addPolygons(
    data = ward_boundaries_1,
    weight = 1,
    opacity = 1.0,
    fillOpacity = 0.6,
    smoothFactor = 0.5,
    color = ~ pal(colorData)
  ) %>% # add labels for ward name
  addLabelOnlyMarkers(
    data = centers,
    lng = ~ x,
    lat = ~ y,
    label = ~ region,
    labelOptions = labelOptions(
      noHide = TRUE,
      direction = 'top',
      textOnly = TRUE,
      textsize = "15px"
    )
  )
map_1
```

# map all service requests across the city

```{r}
map_all_requests <- map_1 %>%
  # add points
  addCircleMarkers(
    data = final_data_sf,
    
    # create custom labels
    label = paste(
      "Address: ", final_data_sf$addr, "<br>",
      "Service request: ",
      final_data_sf$Service.Request.Description,"<br>",
      "Date: ", final_data_sf$Created.Date
    ) %>%
      lapply(htmltools::HTML)
  ) %>%
  # add a legend
  addLegend(
    colors = "blue",
    labels = "Service Request",
    title = "Serive Request Distribution",
    opacity = 1, 
    position = "bottomleft"
  )

map_all_requests
```

# map all service requests as clusters for better visualization and add key places in the city 

```{r}
# create custom icons for key places in the city
icon.fa <- makeAwesomeIcon(
  icon = "flag", markerColor = "red",
  library = "fa",
  iconColor = "black"
)

map_2 <- map_1 %>%
  # add points
  addCircleMarkers(data = final_data_sf,
                   clusterOptions = markerClusterOptions()) %>%
  # add key places to map
  addAwesomeMarkers(
    lat = 42.3188,
    lng = -83.0420,
    label = "St.Clair Centre for the Arts",
    icon = icon.fa
  ) %>% 
  addAwesomeMarkers(
    lat = 42.3043,
    lng = -83.0660,
    label = "University of windsor",
    icon = icon.fa
  ) %>%  
  addAwesomeMarkers(
    lat = 42.2485,
    lng = -83.0203,
    label = "St. Clair College - Main Building",
    icon = icon.fa
  ) %>% 
  addAwesomeMarkers(
    lat = 42.2660,
    lng = -82.9607,
    label = "Windsor International Airport",
    icon = icon.fa
  ) %>% 
  addAwesomeMarkers(
    lat = 42.2660,
    lng = -82.9607,
    label = "Windsor International Airport",
    icon = icon.fa
  ) %>% 
  addAwesomeMarkers(
    lat = 42.3209,
    lng = -83.0183,
    label = "Wyandotte Town Centre",
    icon = icon.fa
  ) %>% 
  addAwesomeMarkers(
    lat = 42.2732,
    lng = -83.0014,
    label = "Devonshire mall",
    icon = icon.fa
  ) %>% 
  addAwesomeMarkers(
    lat = 42.3192,
    lng = -82.9278,
    label = "WFCU Centre",
    icon = icon.fa
  ) %>% addAwesomeMarkers(
    lat = 42.3386,
    lng = -82.9191,
    label = "Sand Point Beach",
    icon = icon.fa
  ) %>% addAwesomeMarkers(
    lat = 42.299961,
    lng = -82.997269,
    label = "Windsor Regional Hospital - Metropolitan Campus",
    icon = icon.fa
  )%>% addAwesomeMarkers(
    lat = 42.2951,
    lng = -83.0227,
    label = "Jackson Park",
    icon = icon.fa
  ) %>% addAwesomeMarkers(
    lat = 42.314070,
    lng = -82.943030,
    label = "Walmart Supercentre",
    icon = icon.fa
  ) %>% addAwesomeMarkers(
    lat = 42.3044,
    lng = -82.9786,
    label = "Central Mall",
    icon = icon.fa
  ) %>% addAwesomeMarkers(
    lat = 42.3044,
    lng = -82.9786,
    label = "Central Mall",
    icon = icon.fa
  ) %>% addAwesomeMarkers(
    lat = 42.2551,
    lng = -82.9656,
    label = "Costco Wholesale",
    icon = icon.fa
  ) %>% addAwesomeMarkers(
    lat = 42.28793019457313,
    lng = -83.0239635485438,
    label = "Service Ontario",
    icon = icon.fa
  )
  

map_2
```

# Contour plot for all service requests 2022

```{r}
# MAKE CONTOUR LINES
kde <- bkde2D(cbind(final_data_2022$longitude,final_data_2022$latitude),
              bandwidth=c(.0045, .0068), gridsize = c(1000,1000))

CL <- contourLines(kde$x1 , kde$x2 , kde$fhat)

# EXTRACT CONTOUR LINE LEVELS
LEVS <- as.factor(sapply(CL, `[[`, "level"))
NLEV <- length(levels(LEVS))

# CONVERT CONTOUR LINES TO POLYGONS
pgons <- lapply(1:length(CL), function(i)
    Polygons(list(Polygon(cbind(CL[[i]]$x, CL[[i]]$y))), ID=i))
spgons = SpatialPolygons(pgons)

# Leaflet map with polygons
leaflet(spgons) %>% addTiles() %>% 
    addPolygons(color = heat.colors(NLEV, NULL)[LEVS])

```

# Create Raster from Kernel Density output

```{r}
KernelDensityRaster <- raster(list(x=kde$x1 ,y=kde$x2 ,z = kde$fhat))

#set low density cells as NA so we can make them transparent with the colorNumeric function
 KernelDensityRaster@data@values[which(KernelDensityRaster@data@values < 1)] <- NA

#create pal function for coloring the raster
palRaster <- colorBin("Spectral", bins = 7, domain = KernelDensityRaster@data@values, na.color = "transparent")

# Leaflet map with raster
leaflet()  %>% addTiles() %>% 
  addRasterImage(KernelDensityRaster, 
                 colors = palRaster, 
                 opacity = .5) %>%
  addLegend(pal = palRaster, 
            values = KernelDensityRaster@data@values, 
            title = "Kernel Density of Points")
```

# Create Raster from Kernel Density output (dark background)

```{r}
KernelDensityRaster <- raster(list(x=kde$x1 ,y=kde$x2 ,z = kde$fhat))

#set low density cells as NA so we can make them transparent with the colorNumeric function
 KernelDensityRaster@data@values[which(KernelDensityRaster@data@values < 1)] <- NA

#create pal function for coloring the raster
palRaster <- colorBin("Spectral", bins = 7, domain = KernelDensityRaster@data@values, na.color = "transparent")

# Leaflet map with raster
leaflet() %>%  addProviderTiles("CartoDB.DarkMatter") %>% 
  addRasterImage(KernelDensityRaster, 
                 colors = palRaster, 
                 opacity = .3) %>%
  addLegend(pal = palRaster, 
            values = KernelDensityRaster@data@values, 
            title = "Kernel Density of Points")

```

# Contour plot for top 5 service requests 2022

## Dirty Yard

```{r}
final_data_2022_dirty_yard <- final_data_2022 %>% filter(Service.Request.Description=='Dirty Yard')

# Create kernel density output
kde <- bkde2D(cbind(final_data_2022_dirty_yard$longitude,final_data_2022_dirty_yard$latitude),
              bandwidth=c(.0045, .0068), gridsize = c(1000,1000))

# create contour lines based on kernel density output
CL <- contourLines(kde$x1 , kde$x2 , kde$fhat)

# extract countour LINE LEVELS
LEVS <- as.factor(sapply(CL, `[[`, "level"))
NLEV <- length(levels(LEVS))

# convert to ratser
KernelDensityRaster <- raster(list(x=kde$x1 ,y=kde$x2 ,z = kde$fhat))

#set low density cells as NA so we can make them transparent with the colorNumeric function
 KernelDensityRaster@data@values[which(KernelDensityRaster@data@values < 1)] <- NA

#create pal function for coloring the raster
palRaster <- colorBin("Spectral", bins = 7, domain = KernelDensityRaster@data@values, na.color = "transparent")

# create raster containing ward labels
map_ward_boundaries %>% 
  addRasterImage(KernelDensityRaster, 
                 colors = palRaster, 
                 opacity = .5) %>%
  addLegend(pal = palRaster, 
            values = KernelDensityRaster@data@values, 
            title = "Kernel Density of Points")

```


## Building Conditions

```{r}
final_data_2022_building_condition <- final_data_2022 %>% filter(Service.Request.Description=='Building Conditions')

# Create kernel density output
kde <- bkde2D(cbind(final_data_2022_building_condition$longitude,final_data_2022_building_condition$latitude),
              bandwidth=c(.0045, .0068), gridsize = c(1000,1000))

# create contour lines based on kernel density output
CL <- contourLines(kde$x1 , kde$x2 , kde$fhat)

# extract countour LINE LEVELS
LEVS <- as.factor(sapply(CL, `[[`, "level"))
NLEV <- length(levels(LEVS))

# convert to raster
KernelDensityRaster <- raster(list(x=kde$x1 ,y=kde$x2 ,z = kde$fhat))

#set low density cells as NA so we can make them transparent with the colorNumeric function
 KernelDensityRaster@data@values[which(KernelDensityRaster@data@values < 1)] <- NA

#create pal function for coloring the raster
palRaster <- colorBin("Spectral", bins = 7, domain = KernelDensityRaster@data@values, na.color = "transparent")

# create raster containing ward labels
map_ward_boundaries %>% 
  addRasterImage(KernelDensityRaster, 
                 colors = palRaster, 
                 opacity = .5) %>%
  addLegend(pal = palRaster, 
            values = KernelDensityRaster@data@values, 
            title = "Kernel Density of Points")

```


## Garbage Not Collected

```{r}
final_data_2022_garbage <- final_data_2022 %>% filter(Service.Request.Description=='Garbage Not Collected')

# Create kernel density output
kde <- bkde2D(cbind(final_data_2022_garbage$longitude,final_data_2022_garbage$latitude),
              bandwidth=c(.0045, .0068), gridsize = c(1000,1000))

# create contour lines based on kernel density output
CL <- contourLines(kde$x1 , kde$x2 , kde$fhat)

# extract countour LINE LEVELS
LEVS <- as.factor(sapply(CL, `[[`, "level"))
NLEV <- length(levels(LEVS))

# convert to ratser
KernelDensityRaster <- raster(list(x=kde$x1 ,y=kde$x2 ,z = kde$fhat))

#set low density cells as NA so we can make them transparent with the colorNumeric function
 KernelDensityRaster@data@values[which(KernelDensityRaster@data@values < 1)] <- NA

#create pal function for coloring the raster
palRaster <- colorBin("Spectral", bins = 7, domain = KernelDensityRaster@data@values, na.color = "transparent")

# create raster containing ward labels
map_ward_boundaries %>% 
  addRasterImage(KernelDensityRaster, 
                 colors = palRaster, 
                 opacity = .5) %>%
  addLegend(pal = palRaster, 
            values = KernelDensityRaster@data@values, 
            title = "Kernel Density of Points")

```


## Rodent Extermination

```{r}
final_data_2022_rodent_extermination <- final_data_2022 %>% filter(Service.Request.Description=='Rodent Extermination')

# Create kernel density output
kde <- bkde2D(cbind(final_data_2022_rodent_extermination$longitude,final_data_2022_rodent_extermination$latitude),
              bandwidth=c(.0045, .0068), gridsize = c(1000,1000))

# create contour lines based on kernel density output
CL <- contourLines(kde$x1 , kde$x2 , kde$fhat)

# extract countour LINE LEVELS
LEVS <- as.factor(sapply(CL, `[[`, "level"))
NLEV <- length(levels(LEVS))

# convert to ratser
KernelDensityRaster <- raster(list(x=kde$x1 ,y=kde$x2 ,z = kde$fhat))

#set low density cells as NA so we can make them transparent with the colorNumeric function
 KernelDensityRaster@data@values[which(KernelDensityRaster@data@values < 1)] <- NA

#create pal function for coloring the raster
palRaster <- colorBin("Spectral", bins = 7, domain = KernelDensityRaster@data@values, na.color = "transparent")

# create raster containing ward labels
map_ward_boundaries %>% 
  addRasterImage(KernelDensityRaster, 
                 colors = palRaster, 
                 opacity = .5) %>%
  addLegend(pal = palRaster, 
            values = KernelDensityRaster@data@values, 
            title = "Kernel Density of Points")

```


## Tree Request

```{r}
final_data_2022_tree_request <- final_data_2022 %>% filter(Service.Request.Description=='Tree Request')

# Create kernel density output
kde <- bkde2D(cbind(final_data_2022_tree_request$longitude,final_data_2022_tree_request$latitude),
              bandwidth=c(.0045, .0068), gridsize = c(1000,1000))

# create contour lines based on kernel density output
CL <- contourLines(kde$x1 , kde$x2 , kde$fhat)

# extract countour LINE LEVELS
LEVS <- as.factor(sapply(CL, `[[`, "level"))
NLEV <- length(levels(LEVS))

# convert to ratser
KernelDensityRaster <- raster(list(x=kde$x1 ,y=kde$x2 ,z = kde$fhat))

#set low density cells as NA so we can make them transparent with the colorNumeric function
 KernelDensityRaster@data@values[which(KernelDensityRaster@data@values < 1)] <- NA

#create pal function for coloring the raster
palRaster <- colorBin("Spectral", bins = 7, domain = KernelDensityRaster@data@values, na.color = "transparent")

# create raster containing ward labels
map_ward_boundaries %>% 
  addRasterImage(KernelDensityRaster, 
                 colors = palRaster, 
                 opacity = .5) %>%
  addLegend(pal = palRaster, 
            values = KernelDensityRaster@data@values, 
            title = "Kernel Density of Points")

```


